---
title: "Assignment-1"
date: "2024-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r message=FALSE}
library(tidyverse)
library(knitr)
library(ggplot2)
```   
#### Import Dataset

```{r}
setwd("/Users/apple/Desktop/8451 Machine Learning for Epi/Assignment-1")
data <- read_csv("./bcdata_Assignment1.csv")
```

#### 1. Construct a summary table 

Construct a table providing summaries of the quantitative features of the dataset including mean, median, minimum value, and maximum value.

```{r}
# Calculate summaries
summary_table <- data.frame(
  Mean = sapply(data, mean),
  Median = sapply(data, median),
  Minimum = sapply(data, min),
  Maximum = sapply(data, max)
)

# Print the summary table
kable(summary_table)
```





#### 2. Recode BMI into the WHO-defined categories 

Recode BMI into the WHO-defined categories below:

* Severely underweight - BMI less than 16.5kg/m^2 

* Underweight - BMI under 18.5 kg/m^2 

* Normal weight - BMI greater than or equal to 18.5 to 24.9 kg/m^2

* Overweight – BMI greater than or equal to 25 to 29.9 kg/m^2 

* Obesity class I – BMI 30 to 34.9 kg/m^2 

* Obesity class II – BMI 35 to 39.9 kg/m^2 

* Obesity class III – BMI greater than or equal to 40 kg/m^2 

```{r}
# Recode BMI into WHO-defined categories
data$BMI_category <- cut(
  data$BMI,
  breaks = c(0, 16.5, 18.5, 25, 30, 35, 40, Inf),
  labels = c(
    "Severely underweight",
    "Underweight",
    "Normal weight",
    "Overweight",
    "Obesity class I",
    "Obesity class II",
    "Obesity class III"
  ),
  include.lowest = TRUE
)

# Check the frequency results of recoded variable
table_result <- table(data$BMI_category)
dimnames(table_result) <- list("BMI category" = names(table_result))
kable(table_result)
```

#### 3. Create a bar chart 

Create a bar chart showing the proportion of breast cancer cases and controls within each BMI category. (That is, the graph should clearly show, what % of each BMI category are cases and what % are controls).

```{r}
# Create a data frame with the proportions
proportions_data <- data.frame(
  BMI_category = data$BMI_category,
  Classification = data$Classification
)

# Calculate proportions
proportions_data <- proportions_data %>%
  group_by(BMI_category, Classification) %>%
  summarise(Count = n()) %>%
  group_by(BMI_category) %>%
  mutate(Proportion = Count / sum(Count) * 100)

# Create a bar chart
ggplot(proportions_data, aes(x = BMI_category, y = Proportion, fill = Classification)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f%%", Proportion), color = "pink"), 
            position = position_stack(vjust = 0.5), 
            size = 4, show.legend = FALSE) +  
  labs(title = "Proportion of Breast Cancer Cases and Controls by BMI Category",
       x = "BMI Category",
       y = "Proportion (%)",
       fill = "Classification") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal()
```

**Bar chart results:**

* Within the underweight category, 100% are controls.

* Within the normal weight category, 56.4% are cases and 43.6% are controls.

* Within the overweight category, 62.2% are cases and 37.8% are controls.

* Within the obesity class I category, 44.4% are cases and 55.6% are controls.

* Within the obesity class II category, 75.0% are cases and 25.0% are controls.

<br>

#### 4. Construct a logistic regression model 

Construct a logistic regression model using breast cancer classification as the outcome and glucose, HOMA-IR, leptin, BMI (continuous) and age as the independent variables.

```{r}
# Recode 'Classification' to numeric values (0 and 1)
data$Classification_numeric <- ifelse(data$Classification == 2, 1, 0)

# Construct logistic regression model
logistic_model <- glm(Classification_numeric ~ Glucose + HOMA + Leptin + BMI + Age, 
                      data = data, 
                      family = "binomial")

# Summarize the model
summary(logistic_model)
cbind(estimate = coef(logistic_model), confint(logistic_model)) |> knitr::kable(digits = 4)
```


**Logistic regression result:**

A 1-unit change in HOMA-IR is associated with the beta estimate of 0.2739 and a 95% confidence interval of (-0.0256, 0.65320).

<br>

#### 5. Construct a linear regression model 

Construct a linear regression model using insulin as the outcome and BMI (continuous), age, and glucose as the independent variables.

```{r}
# Construct linear regression model
linear_model <- lm(Insulin ~ BMI + Age + Glucose, data = data)

# Summarize the model
summary(linear_model)

# Check beta coefficient and 95% CI
cbind(estimate = coef(linear_model), confint(linear_model)) |> knitr::kable(digits = 4)
```

**Linear regression result:**

A 1-unit change in age is associated with the beta estimate of -0.0540 and a 95% confidence interval of (-0.1569, 0.0489).
